---
title: "Mechanistic model SEM interactions"
author: "Xin Jing"
date: "11/30/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# load library
library(tidyverse)
library(lavaan)

# load data
semdata <- readRDS("./data/fundiv_resdata_20200706.rds")

# clean data
# row ids in dry sites
idd <- c(which(grepl("SPA", semdata$PlotID)),
         which(grepl("ITA", semdata$PlotID)))
# generate interactions
semdata <- semdata %>% 
  mutate(light.int = LAI * HLI,
         water.int = RBdeep * water,
         nut.int = RBdeep * STN,
         nut.int2 = RBdeep * SCN,
         LAI.RB.SCN.int = LAI/RBdeep * SCN) %>% 
  mutate(light.int = scale(light.int),
         water.int = scale(water.int),
         nut.int = scale(nut.int),
         nut.int2 = scale(nut.int2),
         LAI.RB.SCN.int = scale(LAI.RB.SCN.int))

```

# Productivity models

1. Productivity + soil CN ratio  
1.1. All sites
```{r, eval=TRUE, echo=TRUE, fig.height=4.5}
# model formulas
prodDeepLN <- '
lnAWP ~ lnBA + SR + LAI + HLI + light.int + RBdeep + water + water.int + SCN + nut.int2
LAI ~ SR + lnBA
RBdeep ~ SR + lnBA
'
fit1.1 <- sem(prodDeepLN, data = semdata)
# summary(fit1.1, standardized = T, rsq = T, fit.measures = F)
fitMeasures(fit1.1, c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"))
semPlot::semPaths(fit1.1, "std",
                  layout = "tree",
                  title = FALSE, curvePivot = TRUE,
                  intercepts = FALSE, residuals = FALSE, thresholds = 0.05,
                   edge.label.cex = 1.2,
                  posCol = c("blue", "red"),
                  nodeLabels = c("AWP", "LAI", "Root \nbiomass",
                                 "Basal \narea", "SR", "HLI",
                                 "LAI\n:HLI", "Root\n:Water", "Water",
                                 "SCN", "Root\n:SCN"),
                  sizeMan = 10)
```

1.2. Dry sites
```{r, eval=TRUE, echo=TRUE, fig.height=4.5}
fit1.2 <- update(fit1.1, data = semdata[idd, ])
# summary(fit1.2, standardized = T, rsq = T, fit.measures = F)
fitMeasures(fit1.2, c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"))
semPlot::semPaths(fit1.2, "std",
                  layout = "tree",
                  title = FALSE, curvePivot = TRUE,
                  intercepts = FALSE, residuals = FALSE, thresholds = 0.05,
                   edge.label.cex = 1.2,
                  posCol = c("blue", "red"),
                  nodeLabels = c("AWP", "LAI", "Root \nbiomass",
                                 "Basal \narea", "SR", "HLI",
                                 "LAI\n:HLI", "Root\n:Water", "Water",
                                 "SCN", "Root\n:SCN"),
                  sizeMan = 10)
```

1.3. Non-dry sites
```{r, eval=TRUE, echo=TRUE, fig.height=4.5}
fit1.3 <- update(fit1.1, data = semdata[-idd, ])
# summary(fit1.3, standardized = T, rsq = T, fit.measures = F)
fitMeasures(fit1.3, c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"))
semPlot::semPaths(fit1.3, "std",
                  layout = "tree",
                  title = FALSE, curvePivot = TRUE,
                  intercepts = FALSE, residuals = FALSE, thresholds = 0.05,
                   edge.label.cex = 1.2,
                  posCol = c("blue", "red"),
                  nodeLabels = c("AWP", "LAI", "Root \nbiomass",
                                 "Basal \narea", "SR", "HLI",
                                 "LAI\n:HLI", "Root\n:Water", "Water",
                                 "SCN", "Root\n:SCN"),
                  sizeMan = 10)
```

\newpage
# Stability models

2. Stability + soil CN ratio  
2.1. All sites
```{r, eval=TRUE, echo=TRUE, fig.height=4.5}
# model formulas
prodDeepLN <- '
lnstab ~ lnBA + SR + LAI + HLI + light.int + RBdeep + water + water.int + SCN + nut.int2
LAI ~ SR + lnBA
RBdeep ~ SR + lnBA
'
fit2.1 <- sem(prodDeepLN, data = semdata)
# summary(fit2.1, standardized = T, rsq = T, fit.measures = F)
fitMeasures(fit2.1, c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"))
semPlot::semPaths(fit2.1, "std",
                  layout = "tree",
                  title = FALSE, curvePivot = TRUE,
                  intercepts = FALSE, residuals = FALSE, thresholds = 0.05,
                   edge.label.cex = 1.2,
                  posCol = c("blue", "red"),
                  nodeLabels = c("Stab", "LAI", "Root \nbiomass",
                                 "Basal \narea", "SR", "HLI",
                                 "LAI\n:HLI", "Root\n:Water", "Water",
                                 "SCN", "Root\n:SCN"),
                  sizeMan = 10)
```

2.2. Dry sites
```{r, eval=TRUE, echo=TRUE, fig.height=4.5}
fit2.2 <- update(fit2.1, data = semdata[idd, ])
# summary(fit2.2, standardized = T, rsq = T, fit.measures = F)
fitMeasures(fit2.2, c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"))
semPlot::semPaths(fit2.2, "std",
                  layout = "tree",
                  title = FALSE, curvePivot = TRUE,
                  intercepts = FALSE, residuals = FALSE, thresholds = 0.05,
                   edge.label.cex = 1.2,
                  posCol = c("blue", "red"),
                  nodeLabels = c("Stab", "LAI", "Root \nbiomass",
                                 "Basal \narea", "SR", "HLI",
                                 "LAI\n:HLI", "Root\n:Water", "Water",
                                 "SCN", "Root\n:SCN"),
                  sizeMan = 10)
```

2.3. Non-dry sites
```{r, eval=TRUE, echo=TRUE, fig.height=4.5}
fit2.3 <- update(fit2.1, data = semdata[-idd, ])
# summary(fit2.3, standardized = T, rsq = T, fit.measures = F)
fitMeasures(fit2.3, c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"))
semPlot::semPaths(fit2.3, "std",
                  layout = "tree",
                  title = FALSE, curvePivot = TRUE,
                  intercepts = FALSE, residuals = FALSE, thresholds = 0.05,
                   edge.label.cex = 1.2,
                  posCol = c("blue", "red"),
                  nodeLabels = c("Stab", "LAI", "Root \nbiomass",
                                 "Basal \narea", "SR", "HLI",
                                 "LAI\n:HLI", "Root\n:Water", "Water",
                                 "SCN", "Root\n:SCN"),
                  sizeMan = 10)
```

\newpage
# Productivity + Stability models

3. Productivity + Stability + soil CN ratio  
3.1. All sites
```{r, echo=TRUE, fig.height=4.5}
# model formulas
prodDeepLN <- '
lnAWP ~ lnBA + SR + LAI + HLI + light.int + RBdeep + water + water.int + SCN + nut.int2
lnstab ~ lnBA + SR + LAI + HLI + light.int + RBdeep + water + water.int + SCN + nut.int2
LAI ~ SR + lnBA
RBdeep ~ SR + lnBA
lnAWP ~~ lnstab
'
fit3.1 <- sem(prodDeepLN, data = semdata)
summary(fit3.1, standardized = T, rsq = T, fit.measures = F)
fitMeasures(fit3.1, c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"))
semPlot::semPaths(fit3.1, "std",
                  layout = "tree",
                  title = FALSE, curvePivot = TRUE,
                  intercepts = FALSE, residuals = FALSE, thresholds = 0.05,
                   edge.label.cex = 1.2,
                  posCol = c("blue", "red"),
                  nodeLabels = c("AWP", "Stab", "LAI", "Root \nbiomass",
                                 "Basal \narea", "SR", "HLI",
                                 "LAI\n:HLI", "Root\n:Water", "Water",
                                 "SCN", "Root\n:SCN"),
                  sizeMan = 10)
```

3.2. Dry sites
```{r, echo=TRUE, fig.height=4.5}
fit3.2 <- update(fit3.1, data = semdata[idd, ])
summary(fit3.2, standardized = T, rsq = T, fit.measures = F)
fitMeasures(fit3.2, c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"))
semPlot::semPaths(fit3.2, "std",
                  layout = "tree",
                  title = FALSE, curvePivot = TRUE,
                  intercepts = FALSE, residuals = FALSE, thresholds = 0.05,
                   edge.label.cex = 1.2,
                  posCol = c("blue", "red"),
                  nodeLabels = c("AWP", "Stab", "LAI", "Root \nbiomass",
                                 "Basal \narea", "SR", "HLI",
                                 "LAI\n:HLI", "Root\n:Water", "Water",
                                 "SCN", "Root\n:SCN"),
                  sizeMan = 10)
```

3.3. Non-dry sites
```{r, echo=TRUE, fig.height=4.5}
fit3.3 <- update(fit3.1, data = semdata[-idd, ])
summary(fit3.3, standardized = T, rsq = T, fit.measures = F)
fitMeasures(fit3.3, c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"))
semPlot::semPaths(fit3.3, "std",
                  layout = "tree",
                  title = FALSE, curvePivot = TRUE,
                  intercepts = FALSE, residuals = FALSE, thresholds = 0.05,
                   edge.label.cex = 1.2,
                  posCol = c("blue", "red"),
                  nodeLabels = c("AWP", "Stab", "LAI", "Root \nbiomass",
                                 "Basal \narea", "SR", "HLI",
                                 "LAI\n:HLI", "Root\n:Water", "Water",
                                 "SCN", "Root\n:SCN"),
                  sizeMan = 10)
```

\newpage
# Side analysis
# 4. Considering LAI/fine root biomass (proxy for above:belowground biomass) x C:N ratio as interaction effect 

4.1. Productivity + LAI/RB ratio 
```{r, echo=TRUE, fig.height=4.5}
# model formulas
prodDeepLN <- '
lnAWP ~ lnBA + SR + LAI.RB.SCN.int
'
fit4.1 <- sem(prodDeepLN, data = semdata)
summary(fit4.1, standardized = T, rsq = T, fit.measures = F)
fitMeasures(fit4.1, c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"))
semPlot::semPaths(fit4.1, "std",
                  layout = "tree",
                  title = FALSE, curvePivot = TRUE,
                  intercepts = FALSE, residuals = FALSE, thresholds = 0.05,
                   edge.label.cex = 1.2,
                  posCol = c("blue", "red"),
                  # nodeLabels = c("AWP", "Stab", "LAI", "Root \nbiomass",
                  #                "Basal \narea", "SR", "HLI",
                  #                "LAI\n:HLI", "Root\n:Water", "Water",
                  #                "SCN", "Root\n:SCN"),
                  sizeMan = 10)
```

4.2. Stability + LAI/RB ratio 
```{r, echo=TRUE, fig.height=4.5}
# model formulas
stabDeepLN <- '
lnstab ~ lnBA + SR + LAI.RB.SCN.int
'
fit4.1 <- sem(stabDeepLN, data = semdata)
summary(fit4.1, standardized = T, rsq = T, fit.measures = F)
fitMeasures(fit4.1, c("chisq", "df", "pvalue", "cfi", "rmsea", "srmr"))
semPlot::semPaths(fit4.1, "std",
                  layout = "tree",
                  title = FALSE, curvePivot = TRUE,
                  intercepts = FALSE, residuals = FALSE, thresholds = 0.05,
                   edge.label.cex = 1.2,
                  posCol = c("blue", "red"),
                  # nodeLabels = c("AWP", "Stab", "LAI", "Root \nbiomass",
                  #                "Basal \narea", "SR", "HLI",
                  #                "LAI\n:HLI", "Root\n:Water", "Water",
                  #                "SCN", "Root\n:SCN"),
                  sizeMan = 10)
```
